---
title: 'Bios 6301: Assignment 2'
author: "Lingjun Fu"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---

*(informally) Due Thursday, 17 September, 1:00 PM*

50 points total.

This assignment won't be submitted until we've covered Rmarkdown.
Create R chunks for each question and insert your R code appropriately.
Check your output by using the `Knit PDF` button in RStudio.

1. **Working with data** In the `datasets` folder on the course GitHub repo, you will find a file called `cancer.csv`, which is a dataset in comma-separated values (csv) format. This is a large cancer incidence dataset that summarizes the incidence of different cancers for various subgroups. (18 points)

    1. Load the data set into R and make it a data frame called `cancer.df`. (2 points)
```{r}
library(RCurl)
x <- getURL("https://raw.githubusercontent.com/fonnesbeck/Bios6301/master/datasets/cancer.csv")
cancer.df=read.csv(text = x)
```
    2. Determine the number of rows and columns in the data frame. (2)
```{r}
nrow(cancer.df)
ncol(cancer.df)
```
    3. Extract the names of the columns in `cancer.df`. (2)
```{r}
colnames(cancer.df)
```
    4. Report the value of the 3000th row in column 6. (2)
```{r}
cancer.df[3000,6]
```
    5. Report the contents of the 172nd row. (2)
```{r}
cancer.df[172,]
```
    6. Create a new column that is the incidence *rate* (per 100,000) for each row.(3)
```{r}
cancer.df["incidence_rate"] <- NA
cancer.df$incidence_rate <- cancer.df$incidence / 100000
```
    7. How many subgroups (rows) have a zero incidence rate? (2)
```{r}
sum(cancer.df$incidence_rate==0)
```
    8. Find the subgroup with the highest incidence rate.(3)
```{r}
cancer.df[which.max(cancer.df$incidence_rate),]
```
2. **Data types** (10 points)

    1. Create the following vector: `x <- c("5","12","7")`. Which of the following commands will produce an error message? For each command, Either explain why they should be errors, or explain the non-erroneous result. (4 points)

            max(x): "7"
            sort(x): "12" "5"  "7"
            sum(x): x is a chracter, while sum needs arguments as numeric or complex or logical vectors.

    2. For the next two commands, either explain their results, or why they should produce errors. (3 points)

            y <- c("5",7,12): "5"  "7"  "12"
            y[2] + y[3]: the type of elements in y is chracter, which cannot be added.

    3. For the next two commands, either explain their results, or why they should produce errors. (3 points)

            z <- data.frame(z1="5",z2=7,z3=12): a data frame with "z1", "z2", "z3" as the column name and 5, 7, 12 as the first row
            z[1,2] + z[1,3]: 19

3. **Data structures** Give R expressions that return the following matrices and vectors (*i.e.* do not construct them manually). (3 points each, 12 total)

    1. $(1,2,3,4,5,6,7,8,7,6,5,4,3,2,1)$
```{r}
c(1:8,7:1)
```    
    2. $(1,2,2,3,3,3,4,4,4,4,5,5,5,5,5)$
```{r}
rep(1:5, 1:5)
```  
    3. $\begin{pmatrix}
      0 & 1 & 1  \\
      1 & 0 & 1  \\
      1 & 1 & 0  \\
    \end{pmatrix}$
```{r}
matrix(1,3,3)-diag(3)
```
    4. $\begin{pmatrix}
      1 & 2 & 3 & 4 \\
      1 & 4 & 9 & 16 \\
      1 & 8 & 27 & 64  \\
      1 & 16 & 81 & 256 \\
      1 & 32 & 243 & 1024  \\
    \end{pmatrix}$
```{r}
pivot<-c(1:4)
matrix(c(pivot,pivot^2,pivot^3,pivot^4,pivot^5), nrow = 5, ncol = 4, byrow = T)
```
4. **Basic programming** (10 points)

    1. Let $h(x,n)=1+x+x^2+\ldots+x^n = \sum_{i=0}^n x^i$. Write an R program to calculate $h(x,n)$ using a `for` loop. (5 points)
```{r}
h = function(x, n){
    sum = 0
    for (i in 0:n){
        sum = sum + x^i
    }
    return(sum)
}
```
    1. If we list all the natural numbers below 10 that are multiples of 3 or 5, we get 3, 5, 6 and 9. The sum of these multiples is 23. Write an R program to perform the following calculations. (5 points)

        1. Find the sum of all the multiples of 3 or 5 below 1,000. (3, [euler1])
```{r}
sum_3or5 = function(x = 1000){
    sum = 0
    for (i in 1:x-1){
        if(i%%3 == 0 | i%%5 == 0)
        sum = sum + i
    }
    return(sum)
}
sum_3or5()
```

        1. Find the sum of all the multiples of 4 or 7 below 1,000,000. (2)
```{r}
sum_4or7 = function(x = 1000000){
    sum = 0
    for (i in 1:x-1){
        if(i%%4 == 0 | i%%7 == 0)
        sum = sum + i
    }
    return(sum)
}
sum_4or7()
```

    1. Each new term in the Fibonacci sequence is generated by adding the previous two terms. By starting with 1 and 2, the first 10 terms will be $(1, 2, 3, 5, 8, 13, 21, 34, 55, 89)$. Write an R program to calculate the sum of the first 15 even-valued terms. (5 bonus points, [euler2])
```{r}
sum_Fibonacci = function(n = 15){
    f1 = 1
    f2 = 2
    count = 1
    sum = 2
    while (count < n){
        temp1 = f1
        temp2 = f2
        f1 = temp2
        f2 = temp1 + temp2
        if(f2%%2 == 0){
            count = count + 1
            sum = sum + f2
        }
    }
    return(sum)
}
sum_Fibonacci()
```
Some problems taken or inspired by projecteuler.

[euler1]:https://projecteuler.net/problem=1
[euler2]: https://projecteuler.net/problem=2
