---
title: 'Bios 6301: Assignment 4'
author: "Lingjun Fu"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---

*Due Tuesday, 27 October, 1:00 PM*

$5^{n=day}$ points taken off for each day late.

50 points total.

Submit a single knitr file (named `homework4.rmd`), along with a valid PDF output file. Inside the file, clearly indicate which parts of your responses go with which problems (you may use the original homework document as a template). Add your name as `author` to the file's metadata section. Raw R code/output or word processor files are not acceptable.

Failure to name file `homework4.rmd` or include author name may result in 5 points taken off.

### Question 1 ###

**15 points**

A problem with the Newton-Raphson algorithm is that it needs the derivative $f'$. If the derivative is hard to compute or does not exist, then we can use the *secant method*, which only requires that the function $f$ is continuous.

Like the Newton-Raphson method, the **secant method** is based on a linear approximation to the function $f$. Suppose that $f$ has a root at $a$. For this method we assume that we have *two* current guesses, $x_0$ and $x_1$, for the value of $a$. We will think of $x_0$ as an older guess and we want to replace the pair $x_0$, $x_1$ by the pair $x_1$, $x_2$, where $x_2$ is a new guess.

To find a good new guess x2 we first draw the straight line from $(x_0,f(x_0))$ to $(x_1,f(x_1))$, which is called a secant of the curve $y = f(x)$. Like the tangent, the secant is a linear approximation of the behavior of $y = f(x)$, in the region of the points $x_0$ and $x_1$. As the new guess we will use the x-coordinate $x_2$ of the point at which the secant crosses the x-axis.

The general form of the recurrence equation for the secant method is:

$$x_{i+1} = x_i - f(x_i)\frac{x_i - x_{i-1}}{f(x_i) - f(x_{i-1})}$$

Notice that we no longer need to know $f'$ but in return we have to provide *two* initial points, $x_0$ and $x_1$.

**Write a function that implements the secant algorithm.** Validate your program by finding the root of the function $f(x) = \cos(x) - x$. Compare its performance with the Newton-Raphson method -- which is faster, and by how much?  For this example $f'(x) = -\sin(x) - 1$.

```{r}
test_fun  <- function(x) cos(x) - x   ### can be any function
test_fun_d <- function(x) -sin(x) - 1   ### the derivative of test function
secant <- function(FUN, x_0, x_1){
    threshold <- 1e-5
    while (abs(x_0 - x_1) > threshold){
        older = x_0
        old = x_1
        x_0 = old
        x_1 = old - FUN(old)*(old - older)/(FUN(old) - FUN(older))
    }
    return (x_1)
}

newton <- function(F1, F2, guess){
    threshold <- 1e-5
    x_0 = guess
    x_1 = guess - F1(guess)/F2(guess)
    while(abs(x_0 - x_1) > threshold){
        x_0 = x_1
        x_1 = x_1 - F1(x_1)/F2(x_1)     
    }
    return (x_1)   
}
n = 1e5
system.time(replicate(n, newton(test_fun, test_fun_d, 0)))
system.time(replicate(n, secant(test_fun, 0, 3.14/2)))
Q1 <- secant(test_fun, 0, 3.14/2) # make a reasonable guess of ranges
```

Using Mathematica FindRoot function, we get 0.739085, which agree with what we get here using both methods. Using system.time function, we find that the Newton-Raphson method is faster than the secant method by about 50%. This makes sense because Newton method is at the compromise of one more input by hand: derivative (everything comes at a price!).

A more detailed discussion can be found at Wikipedia. The secant method can be interpreted as a method in which the derivative is replaced by an approximation and is thus a Quasi-Newton method. If we compare Newton's method with the secant method, we see that Newton's method converges faster. However, Newton's method requires the evaluation of both $f$ and its derivative $f^{\prime}$ at every step, while the secant method only requires the evaluation of $f$. Therefore, the secant method may occasionally be faster in practice. For instance, if we assume that evaluating $f$ takes as much time as evaluating its derivative and we neglect all other costs, we can do two steps of the secant method for the same cost as one step of Newton's method, so the secant method is faster. If however we consider parallel processing for the evaluation of the derivative, Newton's method proves its worth, being faster in time, though still spending more steps.

### Question 2 ###

**18 points**

The game of craps is played as follows. First, you roll two six-sided dice; let x be the sum of the dice on the first roll. If x = 7 or 11 you win, otherwise you keep rolling until either you get x again, in which case you also win, or until you get a 7 or 11, in which case you lose.

Write a program to simulate a game of craps. You can use the following snippet of code to simulate the roll of two (fair) dice:

```{r}
Craps <- function(roll1=NA){
    temp  <-  sum(ceiling(6*runif(2)))
    if(is.na(roll1)){
        roll1 = temp
        if(roll1 == 7 || roll1 == 11){
            return ("win")
        }
        else{
            return (Craps(roll1))
        }
    }
    else{
        if(temp == 7 | temp == 11){
            return ("lose")
        }
        else if(temp == roll1){
            return ("win")
        }
        else{
            return (Craps(roll1))
        }
    }
}
```

1. The instructor should be able to easily import and run your program (function), and obtain output that clearly shows how the game progressed.  Set the RNG seed with `set.seed(100)` and show the output of three games. (lucky 13 points)

```{r}
set.seed(100)
Craps()
Craps()
Craps()
```

1. Find a seed that will win ten straight games.  Consider adding an argument to your function that disables output. Show the output of the ten games. (5 points)

```{r}
res <- rep(NA, 10)
for(i in seq(1:1000)){
    set.seed(i)
    for(j in 1:10){
        res[j] <- Craps()
    }
    if('lose' %in% res == FALSE){
        print (i)
    }   
}
```

The seed is 880.

### Question 3 ###

**12 points**

Obtain a copy of the [football-values lecture](https://github.com/couthcommander/football-values).  Save the five 2015 CSV files in your working directory.

Modify the code to create a function.  This function will create dollar values given information (as arguments) about a league setup.  It will return a data.frame and write this data.frame to a CSV file.  The final data.frame should contain the columns 'PlayerName', 'pos', 'points', 'value' and be orderd by value descendingly.  Do not round dollar values.

Note that the returned data.frame should have `sum(posReq)*nTeams` rows.

Define the function as such (6 points):

```{r}
# download files
library(RCurl)
k15 <-"https://raw.githubusercontent.com/couthcommander/football-values/master/2015/proj_k15.csv"
qb15  <- "https://raw.githubusercontent.com/couthcommander/football-values/master/2015/proj_qb15.csv"
rb15 <- "https://raw.githubusercontent.com/couthcommander/football-values/master/2015/proj_rb15.csv"
te15 <- "https://raw.githubusercontent.com/couthcommander/football-values/master/2015/proj_te15.csv"
wr15 <- "https://raw.githubusercontent.com/couthcommander/football-values/master/2015/proj_wr15.csv"

download.file(k15, destfile="proj_k15.csv",method="curl")
download.file(qb15, destfile="proj_qb15.csv",method="curl")
download.file(rb15, destfile="proj_rb15.csv",method="curl")
download.file(te15, destfile="proj_te15.csv",method="curl")
download.file(wr15, destfile="proj_wr15.csv",method="curl")

# path: directory path to input files
# nTeams: number of teams in league
# cap: money available to each team
# posReq: number of starters for each position
# points: point allocation for each category
ffvalues <- function(path, file='outfile.csv', nTeams=12, cap=200,  
posReq=c(qb=1, rb=2, wr=3, te=1, k=1),  
points=c(fg=4, xpt=1, pass_yds=1/25, pass_tds=4, pass_ints=-2,  
rush_yds=1/10, rush_tds=6, fumbles=-2, rec_yds=1/20, rec_tds=6)){
    setwd(path)
  ## read in CSV files
    k <- read.csv("proj_k15.csv", header=TRUE, stringsAsFactors=FALSE)
    qb <- read.csv("proj_qb15.csv", header=TRUE, stringsAsFactors=FALSE)
    rb <- read.csv("proj_rb15.csv", header=TRUE, stringsAsFactors=FALSE)
    te <- read.csv("proj_te15.csv", header=TRUE, stringsAsFactors=FALSE)
    wr <- read.csv("proj_wr15.csv", header=TRUE, stringsAsFactors=FALSE)
    
  ## calculate dollar values
    cols <- unique(c(names(k), names(qb), names(rb), names(te), names(wr)))
    k[,'pos'] <- 'k'
    qb[,'pos'] <- 'qb'
    rb[,'pos'] <- 'rb'
    te[,'pos'] <- 'te'
    wr[,'pos'] <- 'wr'
    cols <- c(cols, 'pos')
    k[,setdiff(cols, names(k))] <- 0
    qb[,setdiff(cols, names(qb))] <- 0
    rb[,setdiff(cols, names(rb))] <- 0
    te[,setdiff(cols, names(te))] <- 0
    wr[,setdiff(cols, names(wr))] <- 0
    
    x <- rbind(k[,cols], qb[,cols], rb[,cols], te[,cols], wr[,cols])

    x[,'p_fg'] <- x[,'fg']*points[["fg"]]
    x[,'p_xpt'] <- x[,'xpt']*points[["xpt"]]
    x[,'p_pass_yds'] <- x[,'pass_yds']*points[["pass_yds"]]
    x[,'p_pass_tds'] <- x[,'pass_tds']*points[["pass_tds"]]
    x[,'p_pass_ints'] <- x[,'pass_ints']*points[["pass_ints"]]
    x[,'p_rush_yds'] <- x[,'rush_yds']*points[["rush_yds"]]
    x[,'p_rush_tds'] <- x[,'rush_tds']*points[["rush_tds"]]
    x[,'p_fumbles'] <- x[,'fumbles']*points[["fumbles"]]
    x[,'p_rec_yds'] <- x[,'rec_yds']*points[["rec_yds"]]
    x[,'p_rec_tds'] <- x[,'rec_tds']*points[["rec_tds"]]
 
    x[,'points'] <- rowSums(x[,grep("^p_", names(x))])
  
    x2 <- x[order(x[,'points'], decreasing=TRUE),]
    k.ix <- which(x2[,'pos']=='k')
    qb.ix <- which(x2[,'pos']=='qb')
    rb.ix <- which(x2[,'pos']=='rb')
    te.ix <- which(x2[,'pos']=='te')
    wr.ix <- which(x2[,'pos']=='wr')
    
    x2[k.ix, 'marg'] <- x2[k.ix,'points'] - x2[k.ix[nTeams*posReq[["k"]]],'points']  
    x2[qb.ix, 'marg'] <- x2[qb.ix,'points'] - x2[qb.ix[nTeams*posReq[["qb"]]],'points']
    x2[rb.ix, 'marg'] <- x2[rb.ix,'points'] - x2[rb.ix[nTeams*posReq[["rb"]]],'points']
    x2[te.ix, 'marg'] <- x2[te.ix,'points'] - x2[te.ix[nTeams*posReq[["te"]]],'points']
    x2[wr.ix, 'marg'] <- x2[wr.ix,'points'] - x2[wr.ix[nTeams*posReq[["wr"]]],'points']

    x3 <- x2[x2[,'marg'] >= 0,]
    x3 <- x3[order(x3[,'marg'], decreasing=TRUE),]
    rownames(x3) <- NULL
    x3[,'value'] <- x3[,'marg']*(nTeams*cap-nrow(x3))/sum(x3[,'marg']) + 1
    x4 <- x3[,c('PlayerName','pos','points','value')]
    x4 <- x4[order(x4[,'value'], decreasing=TRUE),]
  
  ## save dollar values as CSV file
    write.csv(x4, file = file)
  
  ## return data.frame with dollar values
    return (x4)
}
```

1. Call `x1 <- ffvalues('.')`

```{r}
x1 <- ffvalues('.')
Q1_1 <- sum(x1$value>20)
Q1_2  <- x1[which(x1[,'pos']=='rb')[15],1]
Q1_1
Q1_2
```

    1.  How many players are worth more than $20? (1 point)
    
    Q1_1 = 40
    
    1.  Who is 15th most valuable running back (rb)? (1 point)
    
    Q1_2 = "Melvin Gordon"

1. Call `x2 <- ffvalues(getwd(), '16team.csv', nTeams=16, cap=150)`

```{r}
x2 <- ffvalues(getwd(), '16team.csv', nTeams=16, cap=150)
Q2_1 <- sum(x2$value>20)
y <- x2[1:40,]
Q2_2  <- nrow(y[y[2]=='wr',])
Q2_1
Q2_2
```

    1.  How many players are worth more than $20? (1 point)
    
    Q2_1 = 41
    
    1.  How many wide receivers (wr) are in the top 40? (1 point)
    
    Q2_2 = 13

1. Call:

```{r}
### need slight modification to accommodate with k=0 case 
ffvalues_k_0 <- function(path, file='qbheavy.csv', nTeams=12, cap=200,  
posReq=c(qb=2, rb=2, wr=3, te=1, k=0),
points=c(fg=0, xpt=0, pass_yds=1/25,  
pass_tds=6, pass_ints=-2, rush_yds=1/10,  
rush_tds=6, fumbles=-2, rec_yds=1/20, rec_tds=6)){
    setwd(path)
  ## read in CSV files
    k <- read.csv("proj_k15.csv", header=TRUE, stringsAsFactors=FALSE)
    qb <- read.csv("proj_qb15.csv", header=TRUE, stringsAsFactors=FALSE)
    rb <- read.csv("proj_rb15.csv", header=TRUE, stringsAsFactors=FALSE)
    te <- read.csv("proj_te15.csv", header=TRUE, stringsAsFactors=FALSE)
    wr <- read.csv("proj_wr15.csv", header=TRUE, stringsAsFactors=FALSE)
    
  ## calculate dollar values
    cols <- unique(c(names(k),names(qb), names(rb), names(te), names(wr)))
    k[,'pos'] <- 'k'
    qb[,'pos'] <- 'qb'
    rb[,'pos'] <- 'rb'
    te[,'pos'] <- 'te'
    wr[,'pos'] <- 'wr'
  
    cols <- c(cols, 'pos')
    k[,setdiff(cols, names(k))] <- 0
    qb[,setdiff(cols, names(qb))] <- 0
    rb[,setdiff(cols, names(rb))] <- 0
    te[,setdiff(cols, names(te))] <- 0
    wr[,setdiff(cols, names(wr))] <- 0
    
    x <- rbind(k[,cols], qb[,cols], rb[,cols], te[,cols], wr[,cols])

    x[,'p_fg'] <- x[,'fg']*points[["fg"]]
    x[,'p_xpt'] <- x[,'xpt']*points[["xpt"]]
    x[,'p_pass_yds'] <- x[,'pass_yds']*points[["pass_yds"]]
    x[,'p_pass_tds'] <- x[,'pass_tds']*points[["pass_tds"]]
    x[,'p_pass_ints'] <- x[,'pass_ints']*points[["pass_ints"]]
    x[,'p_rush_yds'] <- x[,'rush_yds']*points[["rush_yds"]]
    x[,'p_rush_tds'] <- x[,'rush_tds']*points[["rush_tds"]]
    x[,'p_fumbles'] <- x[,'fumbles']*points[["fumbles"]]
    x[,'p_rec_yds'] <- x[,'rec_yds']*points[["rec_yds"]]
    x[,'p_rec_tds'] <- x[,'rec_tds']*points[["rec_tds"]]
 
    x[,'points'] <- rowSums(x[,grep("^p_", names(x))])
  
    x2 <- x[order(x[,'points'], decreasing=TRUE),]
    k.ix <- which(x2[,'pos']=='k')
    qb.ix <- which(x2[,'pos']=='qb')
    rb.ix <- which(x2[,'pos']=='rb')
    te.ix <- which(x2[,'pos']=='te')
    wr.ix <- which(x2[,'pos']=='wr')
     
    x2[qb.ix, 'marg'] <- x2[qb.ix,'points'] - x2[qb.ix[max(1,posReq['qb'])*nTeams], 'points']
    x2[rb.ix, 'marg'] <- x2[rb.ix,'points'] - x2[rb.ix[max(1,posReq['rb'])*nTeams], 'points']
    x2[te.ix, 'marg'] <- x2[te.ix,'points'] - x2[te.ix[max(1,posReq['te'])*nTeams], 'points']
    x2[wr.ix, 'marg'] <- x2[wr.ix,'points'] - x2[wr.ix[max(1,posReq['wr'])*nTeams], 'points']
    x2[k.ix, 'marg'] <- x2[k.ix,'points'] - x2[k.ix[max(1,posReq['k'])*nTeams], 'points']

    x3 <- x2[x2[,'marg'] >= 0,]
    x3 <- x3[order(x3[,'marg'], decreasing=TRUE),]
    rownames(x3) <- NULL
    x3[,'value'] <- x3[,'marg']*(nTeams*cap-nrow(x3))/sum(x3[,'marg']) + 1
    x4 <- x3[,c('PlayerName','pos','points','value')]
    x4 <- x4[order(x4[,'value'], decreasing=TRUE),]
  
  ## save dollar values as CSV file
    write.csv(x4, file = file)
  
  ## return data.frame with dollar values
    data.frame(x4)
    return (x4)
}

x3 <- ffvalues_k_0('.', 'qbheavy.csv', posReq=c(qb=2, rb=2, wr=3, te=1, k=0), points=c(fg=0, xpt=0, pass_yds=1/25, pass_tds=6, pass_ints=-2, rush_yds=1/10, rush_tds=6, fumbles=-2, rec_yds=1/20, rec_tds=6))

Q3_1 <- sum(x3[,'value']>20)
z <- x3[1:30,]
Q3_2  <- nrow(z[z[2]=='qb',]) 
Q3_1
Q3_2
```

    1.  How many players are worth more than $20? (1 point)
    
    Q3_1 = 44
    
    1.  How many quarterbacks (qb) are in the top 30? (1 point)
    
    Q3_2 = 13
    
    

### Question 4 ###

**5 points**

This code makes a list of all functions in the base package:

```{r eval=FALSE}
objs <- mget(ls("package:base"), inherits = TRUE)
funs <- Filter(is.function, objs)
```

Using this list, write code to answer these questions.

1. Which function has the most arguments? (3 points)
```{r}
objs <- mget(ls("package:base"), inherits = TRUE)
funs <- Filter(is.function, objs)
which.max(lapply(funs, function(x) length(formals(x))))
```

1. How many functions have no arguments? (2 points)
```{r}
length(funs[lapply(funs, function(x) length(formals(x)))==0])
```

Hint: find a function that returns the arguments for a given function.

---

### Notes/Grade ###
```{r}
# Question 1,2,3,4: awesome!
```

50/50 points